---
title: "Portfolio"
output: html_document
---

# Portfolio

This portfolio shows how to create a SQL database in R. 2 excel datasets(tags, days) are used in SQL databases and they are avialable at GitHub. Data stored in these excel files are collected from Twitter API and includes user and tweet information.

Libraries required for this process are written below:

If they are not installed before install.packages('readxl'), install.packages('dplyr') and  install.packages('RSQLite') codes should be run. RQLite is available on CRAN.

```{r}

library(readxl)
library(dplyr)
library(RSQLite)

```

While first two libraries are for importing excel files and R analysis, open source RSQLiteis the primary library used for managing SQL database.

```{r}
h <-  read_excel("C:/Users/kilic/Desktop/tags.xlsx")
t <- read_excel("C:/Users/kilic/Desktop/days.xlsx")
```

## Pre-processing

SQL is a relational database whose structure enables us to identify and reach data in relation to another piece of data in the database. In order to use SQL, those relationaltables should satisfy at least one of the normal forms: 1NF(First Normal Formal), 2NF(Second Normal Form) and 3NF(Third Normal Form).

In this portfolio three data frames satisfting following rules of 1NF are created initially:

1)Each table has primary key (minimal set if attributes defining unique records). Primary keys uniquely identifies rows/records of the table.
2)Columns store different information in the same table. There are not any repeating columns or rows.
3)1NF leads reduction in data reduncies and can be derived by sing primary keys, avoiding dublicate fiels, repeating groups.

Format of twitter id column is converted to numeric and non-scientific while uniqueness of user_tibble's rows is satisfied with unique() function. Moreover, format of created_at column which consists of date and time information is converted into a right format. After those pre-processing steps 3 data frames names tweet_tibble, user_tibble and hashtag_tibble are ready to be inserted in SQL databases. Each table has its primary key which uniquely defines records and foreign key allowing us to relate/join them.

```{r}
t$id <- as.numeric(t$id)
h$id <- as.numeric(h$id)

t$id <- format(t$id, scientific=F)
h$id <- format(h$id, scientific=F)

t$Time <- format(as.POSIXct(t$created_at,format="%Y:%m:%d %H:%M:%S"),"%H:%M:%S")
t$Date <- format(as.POSIXct(t$created_at,format="%Y:%m:%d %H:%M:%S"),"%Y-%m-%d")

tweet_tibble <- t %>%
  select(id, user_screen_name, retweet_count, text, Date, Time)

user_tibble_d <- t %>%
  select(user_screen_name, user_followers_count, user_location)

user_tibble <- unique(user_tibble_d)

hashtag_tibble <- h 

```

## Creating Database

dbConnect() function of RSQLite is used to open a connection to a database. If the named database does not yet exist, one is created. Database is named as "Table" in this case.
Database will actually exist after some data has been entered in it.

```{r}

db <- dbConnect(SQLite(), dbname="TablePortfolio.sqlite")

```

Data can be inserted both by hand or using imported csv/excel data. First of all, following commands (dbSendQuery()) are used to create UserTable, TweetTable and HashtagTable in the database. In addition, names of the variables/columns, their type, primary/foreign keys are set.

```{r}
#Empty databases are created, keys and type of variables are defined

#User Tables are created 
dbSendQuery(conn = db,  "CREATE TABLE UserTable (
            user_screen_name TEXT PRIMARY KEY,
            user_followers_count INT,
            user_location TEXT)" )

#Tweet Tables are created 

dbSendQuery(conn = db,  "CREATE TABLE TweetTable (
            id INT PRIMARY KEY,
            user_screen_name TEXT,
            favourite_count INT,
            retweet_count INT, 
            urls TEXT,
            media TEXT,
            text TEXT,
            Date DATE,
            Time TIME,
            FOREIGN KEY (user_screen_name) REFERENCES UserTable (user_screen_name) ) ")

#Hashtag Table is created

dbSendQuery(conn = db,  "CREATE TABLE HashtagTable (
            hashtag_id INT PRIMARY KEY,
            id INT,
            hashtags TEXT,
            FOREIGN KEY(id) REFERENCES TWeetTable (id) ) ")

```

Tables have been created but they do not yet contain any data. Thus, inserting data into those tables is the second step. Although, excel files are imported into database in this project, example of inserting data one by one into UserTable is also shown below.

e.g.
dbSendQuery(conn = db,"INSERT INTO UserTable values('ekilic', 30, 'Boston')")

However, inserting each record by hand is not an efficient way especially for huge data. Thus, another way to insert whole data into SQL database in data frame is shown below.


Using dbWriteTable() function, preprocessed excel data (data frames) is entered into those SQL tables.

```{r}


dbWriteTable(conn = db, name = "TweetTable", value = tweet_tibble, overwrite=TRUE)
dbWriteTable(conn = db, name = "UserTable", value = user_tibble, overwrite=TRUE)
dbWriteTable(conn = db, name = "HashtagTable", value = hashtag_tibble,overwrite=TRUE)

```

##Example Query

TweetTable and HashtagTable can be joined by their shared variable-Tweet id, Following code returns users' names who used 'sense8' in their tweets

```{r}
dbGetQuery(db, "SELECT TweetTable.user_screen_name
           FROM TweetTable INNER JOIN HashtagTable ON TweetTable.id == HashtagTable.id
           WHERE HashtagTable.hashtags == 'sense8'")
```


